---
title: "Cryptocurrency Volatility Forecasting"
author: "Elliot Yang"
date: today
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

This project develops an automated cryptocurrency volatility forecasting framework using GARCH-family models.

The workflow consists of:

1.  Downloading daily cryptocurrency prices
2.  Calculating log returns
3.  Testing stationarity using the Augmented Dickey-Fuller (ADF) test
4.  Testing ARCH effects
5.  Estimating GARCH(1,1), GJR-GARCH(1,1), and EGARCH(1,1)
6.  Selecting the best-performing model using AIC, BIC and Log-Likelihood
7.  Forecasting future volatility
8.  Evaluating forecasting performance using RMSE and MAE
9.  Estimating 95% and 99% Value-at-Risk

## Data overview

### Cryptocurrency Returns

```{r}
library(tidyverse)
returns <- read_csv("../data/processed/cleaned_returns.csv")

glimpse(returns)
```

```{r}
head(returns)
```

## Best model

### Selected Models

```{r}
best_models <- readRDS("../outputs/models/best_models.rds")

data.frame(
  Cryptocurrency = names(best_models),
  Model = sapply(best_models,
                 function(x) x@model$modeldesc$vmodel)
)
```

## Forcast accuracy

```{r}
accuracy <- read_csv("../outputs/tables/forecast_accuracy.csv")

knitr::kable(accuracy)
```

## VaR

```{r}
var <- read_csv("../outputs/tables/var_results.csv")

knitr::kable(var)
```

## Forecast plots

```{r}

library(ggplot2)

coins <- c("BTC", "ETH", "SOL", "BNB")

for (coin in coins) {

  forecast_df <- read_csv(
    paste0("../outputs/forecasts/", coin, "_forecast.csv"),
    show_col_types = FALSE
  )

  p <- ggplot(forecast_df,
              aes(x = Date, y = Forecast_Volatility)) +
    geom_line(color = "steelblue", linewidth = 0.8) +
    labs(
      title = paste(coin, "Forecast Volatility"),
      x = "Date",
      y = "Forecast Volatility"
    ) +
    theme_minimal()

  print(p)
}
```
